---
title: "Monarchs Sampling Design - Canada"
author: "Guillaume Larocque"
date: "25/01/2022"
output: html_document
---

```{r setup, eval=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## CREATING THE BASE LAYERS FROM CEC LAND COVER MAP
#library(rgrass7)

#initGRASS('/usr/lib/grass74/',gisDbase = '/home/glaroc/Monarques/GRASS',location = 'CEC',mapset = 'PERMANENT',override = TRUE)
```{r, eval=TRUE}
library(rgrass7)
library(sp)
library(stringr)
library(spatialEco)
library(rgdal)
initGRASS('/usr/lib/grass74/',gisDbase = '/home/ubuntu/data/GRASS',location = 'monarques',mapset = 'PERMANENT',override = TRUE)


buffer=50000
nrand=0.05
use_sp()

```


```{r, eval=FALSE}


##DEFINE THE LIMITS OF THE RANGE BASED ON MONARCHS AND MILKWEED DISTRIBUTION

execGRASS('v.in.ogr',input="/home/ubuntu/data/Monarques/danaus_asclepias_complete_cec.shp",output='range', flags=c('overwrite'))
execGRASS('v.buffer',input="range", distance=buffer, output="range_buffer", flags=c('overwrite'))
execGRASS('v.in.ogr',input='/home/ubuntu/data/Monarques/canada_master_sample_poly_cec.shp', output='gtrs_10', flags=c('overwrite'))
execGRASS('v.in.ogr',input='/home/ubuntu/data/Monarques/canada_mastersample_cec.shp', output='gtrs_pts_10', flags=c('overwrite'))
execGRASS('v.clip',input='gtrs_10',clip='range_buffer',output='gtrs_range', flags=c('overwrite'))

##CREATION OF THE BASE LAND COVER LAYER WITH ROADS, RAILS AND TRANSMISSION LINES
execGRASS('r.in.gdal',input = '/home/ubuntu/data/Monarques/CAN_NALCMS_2015_LC_30m_LAEA_mmu5pix_.tif',output = 'CEC', memory=100000, flags=c('overwrite'))
execGRASS('g.region',raster='CEC')
execGRASS('v.in.ogr',input='/home/ubuntu/data/Monarques/canada_hwy.shp', output= 'Roads', flags=c('overwrite'))
execGRASS('v.in.ogr',input='/home/ubuntu/data/Monarques/canada_rails_cec.shp', output= 'Rails', flags=c('overwrite'))
execGRASS('v.in.ogr',input='/home/ubuntu/data/Monarques/transmission_lines_OSM_CEC.gpkg', output= 'Transmission', flags=c('overwrite'))

execGRASS('v.in.ogr',input='/home/ubuntu/data/Monarques/protected_areas_canada_qc.shp', output= 'Protected', snap=1e-09, flags=c('overwrite'))
execGRASS('v.to.rast',input= 'Roads',output = 'Roads', use='val',value=20, memory=150000, flags=c('overwrite'))
execGRASS('v.to.rast',input= 'Rails',output = 'Rails', use='val',value=21, memory=150000, flags=c('overwrite'))
execGRASS('v.to.rast',input= 'Transmission',output = 'Transmission', use='val',value=22, memory=150000, flags=c('overwrite'))
execGRASS('v.to.rast',input= 'Protected',output = 'Protected', use='val',value=100, memory=150000, flags=c('overwrite'))

execGRASS('r.patch',input='Roads,Rails,Transmission,CEC',output='CEC2', flags=c('overwrite'))
execGRASS('r.null',null=0,map='CEC2')
execGRASS('r.null',null=0,map='Protected')
execGRASS('r.mapcalc',expression='CEC_pa = CEC2+Protected', flags=c('overwrite'))
```

```{r, eval=FALSE}
execGRASS('r.out.gdal', input='CEC_pa',output='/home/ubuntu/data/Monarques/cec_pa_lines.tif', format='GTiff',createopt="COMPRESS=DEFLATE", flags=c('overwrite'))
##LAND COVER SUITABLE FOR MONARCHS
execGRASS('r.reclass', input = 'CEC_pa', output='CEC_suitable', rules='/home/ubuntu/data/Monarques/cec_reclass.txt', flags=c('overwrite'))


##EXTRACT LAND COVERS FOR EACH GRTS ZONE
execGRASS('v.to.rast',input='gtrs_range', output='gtrs_range', use='attr',attribute_column='GRTS_ID',memory=150000, flags=c('overwrite'))
execGRASS('r.stats.zonal',base='gtrs_range', cover='CEC_suitable',method='sum',output='gtrs_suitable_zonal', flags=c('overwrite'))
execGRASS('r.out.gdal',input='gtrs_suitable_zonal',output='/home/ubuntu/data/Monarques/gtrs_suitable_zonal.tif', format='GTiff',createopt="COMPRESS=DEFLATE", flags=c('overwrite'))
execGRASS('r.to.vect',input='gtrs_suitable_zonal', output='gtrs_suitable_zonal',type='area',column='suitable', flags=c('overwrite'))

execGRASS('v.out.ogr',input='gtrs_suitable_zonal',output='/home/ubuntu/data/Monarques/gtrs_suitable_zonal.gpkg',type='area', flags=c('overwrite'))

execGRASS('v.db.addcolumn', map='gtrs_suitable_zonal',col="area double")
execGRASS('v.to.db',map='gtrs_suitable_zonal',option='area',columns='area')
```

```{r, eval=FALSE}
gtrs=readVECT('gtrs_suitable_zonal')
rnd=ceiling(nrow(gtrs)*nrand)
execGRASS('v.extract',input='gtrs_suitable_zonal',output='tmp',where='suitable>50000 AND area>99000000',flags=c('overwrite'))
execGRASS('v.extract',input='tmp',output='selected_blocks',random=rnd,flags=c('overwrite'))
execGRASS('v.db.addcolumn',map='selected_blocks',columns="grts_id int")
execGRASS('v.what.vect',map='selected_blocks',query_map='gtrs_pts_10',query_column='grts_id',column='grts_id')
execGRASS('v.out.ogr',input='selected_blocks', type='area',output='/home/ubuntu/data/Monarques/selected_blocks.gpkg', flags=c('overwrite'))

```

### SELECTION OF GTRS SQUARES ZONES

# Old approach
#gtrs_suitable_zonal file is opened in QGIS and a random identifier is given to each polygon. A random 5% of grid blocks are choosen. 



### 30 M SELECTION
```{r, eval=TRUE}
execGRASS('r.reclass',input='CEC_pa',output='CEC_6cats',rules='/home/ubuntu/data/Monarques/Monarques-repo/cec_reclass_6cats.txt',c('overwrite'))
fids=execGRASS('v.db.select',map='selected_blocks',columns='grts_id',flags=c("c","v"))
fids=attr(fids,"resOut")
ii=0
for (i in fids) {
  execGRASS('v.extract',input='selected_blocks',output='block_temp',where=paste0("grts_id=",i),flags = c('overwrite'))
  execGRASS('g.region', vector='block_temp')
  execGRASS('r.mask',vector='block_temp', flags = c("overwrite"))
  execGRASS('r.to.vect',input='CEC_6cats', output='temp', type='point',flags = c('overwrite'))
  for (j in (1:6)) {
    execGRASS('v.extract',input='temp', output='temp2', where=paste0("value=",j),flags = c('overwrite'))
    tt=execGRASS('v.info',map='temp2',flags=c('t'))
    npoints=as.numeric(str_replace(attr(tt,"resOut")[2],'points=',''))
    if(npoints>1){
      pts <- readVECT('temp2',type='point')
      nsamples <- min(npoints-1,20)
      if(nsamples > 1){
        if(npoints != 20){
          pts2 <- subsample.distance(pts, size=nsamples, d=300)
        }else{
          pts2 <- pts
        }
        writeVECT(pts2, 'temp3',v.in.ogr_flags=c('o','overwrite','t')) 
        if(i==fids[1] && j==1){
          execGRASS('v.patch',input='temp3', output='all30',flags = c('overwrite'))
        }else{
          execGRASS('v.patch',input='temp3', output='all30',flags = c('a','overwrite'))
        }
      }
    }
  }
  ii=ii+1
  print(ii/length(fids))
}

```

```{r, eval=TRUE}
#This doesn't quite work. Used approach below instead. 
execGRASS('g.region',raster='CEC_pa')
#execGRASS('r.mask', flags = c("r","quiet"))
#execGRASS('v.db.droptable',map='all30',flags=c('f'))
#execGRASS('v.db.addtable',map='all30',columns="cover_type int")
#execGRASS('v.what.rast',map='all30',raster='CEC_pa',column='cover_type')
#execGRASS('v.db.addcolumn',map='all30',columns="grts_id int")
#execGRASS('v.what.vect',map='all30',query_map='gtrs_10',query_column='grts_id',column='grts_id')
execGRASS('v.out.ogr', input='all30', type='point', output='/home/ubuntu/data/Monarques/all30.gpkg',flags=c('overwrite'))

```

### GENERATION OF FINAL MAP

Point sampling tool is used in QGIS to combine the province shapefile, the CEC_pa land cover and the GRTS block ids to the attribute table. 

Then, to generate point ID

```{r, eval=FALSE}
execGRASS('v.in.ogr',input='/home/ubuntu/data/Monarques/all30_with_all_cats2.gpkg', output= 'all30_with_cats', flags=c('overwrite','o'))
execGRASS('v.out.ogr', input='all30_with_cats', type='point', output='/home/ubuntu/data/Monarques/all30_with_cats.csv',format='CSV',flags=c('overwrite'))
all30<-read.csv('/home/ubuntu/data/Monarques/all30_with_cats.csv')
for (i in unique(all30$grts_id)) {
  all30[all30$grts_id==i,'point_id'] <- seq(1,sum(all30$grts_id==i))
}
all30$point_id_full=paste0('CA-',str_pad(all30$grts_id,4,side='left',pad=0),'-',all30$point_id)

write.csv(all30,'/home/ubuntu/data/Monarques/all30_with_ids.csv')
```



